A matrix-free approach for finite-strain hyperelastic problems using geometric multigrid

### 2019-01-08 doc/large_strain.tex
remove TBB, expand on how SIMD is done

    In this work, we extend the application of matrix-free approaches to problems in solid mechanics and investigate different implementations of the finite-strain hyperelastic finite element tangent operator numerically.
    In order to improve the convergence behavior of iterative solvers, we also propose a method by which to construct level tangent operators
    and employ them to define a geometric multigrid preconditioner.
    Our implementation employs MPI and Intel Threading Building Blocks parallelization, and SIMD vectorization.
    The performance of the matrix-free operator and the geometric multigrid preconditioner is compared to the matrix-based implementation with an algebraic multigrid preconditioner for a representative numerical example of a heterogeneous hyperelastic material in two and three dimensions.
  \end{abstract}

@@

In this work, we investigate different implementations of the finite-strain hyperelastic tangent operator with respect to their computational cost.
In order to improve the convergence of iterative solvers, we also propose a method by which to construct level matrix-free tangent operators
and employ them to define a \mbox{geometric} multigrid preconditioner.
Our implementation employs MPI and Intel Threading Building Blocks parallelization, together with SIMD vectorization.

@@

We briefly note that the SIMD vectorization is applied at the finite element level, i.e. the matrix-free operator is applied simultaneously on several elements (called ``blocks'') whereas the shared memory parallelization using Intel Threading Building Blocks is adopted at the element blocks.

@@

@@

In this section, we apply the proposed matrix-free operator algorithms for finite-strain hyperelastic materials as well as the geometric multigrid preconditioner to a benchmark problem from \cite{Miehe2007}. The coarse meshes for the 2D and 3D problems are illustrated in Figure \ref{fig:miehe}. The 2D material consists of a square (depicted in blue), a hole and two spherical inclusions (depicted in red). The size of the square domain is $10^{-3} \rm{mm}$.
The matrix material is taken to have Poisson's ratio $0.3$ and shear modulus $\mu=0.4225 \times 10^6\, \rm{N/mm^2}$. The inclusion is taken to be $100$ times stiffer.
The 3D material is obtained by extrusion of the 2D geometry into the third dimension.
Note that although a relatively coarse mesh is used at the coarsest multigrid level, the geometry of the inclusions is captured accurately by manifold descriptions of the boundaries and interfaces.
Both domains are fully fixed along the bottom surface and a distributed load is applied at the top in the $(1,0)$ or $(1,1,0)$ direction for the 2D and the 3D problems, respectively.
The force density $12.5 \times 10^3 \rm{N/mm^2}$ or $12.5 \sqrt{2} \times 10^3 \rm{N/mm^3}$ is applied in 5 steps for the 2D and the 3D problems, respectively.
With respect to the nonlinear solver, the displacement tolerance of the $\mathcal{\ell}_2$ norm for the Newton update is taken to be $10^{-5}$ whereas the relative and absolute tolerances for the residual forces is $10^{-8}$. The relative convergence criteria for the linear solver is $10^{-6}$.
Shared memory parallelization using Intel Threading Building Blocks was disabled for this study.

>>>>

    In this work, we extend the application of matrix-free approaches to problems in solid mechanics and investigate different implementations of the finite-strain hyperelastic finite element tangent operator numerically.
    In order to improve the convergence behavior of iterative solvers, we also propose a method by which to construct level tangent operators
    and employ them to define a geometric multigrid preconditioner.
    The performance of the matrix-free operator and the geometric multigrid preconditioner is compared to the matrix-based implementation with an algebraic multigrid preconditioner for a representative numerical example of a heterogeneous hyperelastic material in two and three dimensions.
  \end{abstract}

@@

In this work, we investigate different implementations of the finite-strain hyperelastic tangent operator with respect to their computational cost.
In order to improve the convergence of iterative solvers, we also propose a method by which to construct level matrix-free tangent operators
and employ them to define a \mbox{geometric} multigrid preconditioner.
MPI parallelization together with SIMD vectorization
{\color{red}
are provided by the \texttt{deal.II} \cite{dealII90} finite element library, used for this study.
}

@@

@@

Note that the SIMD vectorization
{\color{red}
in \texttt{deal.II} \cite{dealII90}
}
is applied at the finite element level, i.e. the matrix-free operator is applied simultaneously on several elements (called ``blocks'').
{\color{red}
For a given quadrature point, we simultaneously perform tensor evaluations and contractions on all elements. This is achieved using templated number types within the \textit{Tensor} class of the \texttt{deal.II} library.
\textit{FEEvaluation} class already provides all the necessary contraction (such as $\overline{\grad^s} \gz N_i : \boldsymbol{\mathcal{C}}\gz g_s$) to implement the three algorithms proposed above.
Thanks to the object oriented design of the \texttt{deal.II} library, the core of those matrix-free algorithms can be implemented in only a handful of lines.
For more information about the implementation details and performance analysis (????), see \cite{kronbichler12}.
}

@@

In this section, we apply the proposed matrix-free operator algorithms for finite-strain hyperelastic materials as well as the geometric multigrid preconditioner to a benchmark problem from \cite{Miehe2007}. The coarse meshes for the 2D and 3D problems are illustrated in Figure \ref{fig:miehe}. The 2D material consists of a square (depicted in blue), a hole and two spherical inclusions (depicted in red). The size of the square domain is $10^{-3} \rm{mm}$.
The matrix material is taken to have Poisson's ratio $0.3$ and shear modulus $\mu=0.4225 \times 10^6\, \rm{N/mm^2}$. The inclusion is taken to be $100$ times stiffer.
The 3D material is obtained by extrusion of the 2D geometry into the third dimension.
Note that although a relatively coarse mesh is used at the coarsest multigrid level, the geometry of the inclusions is captured accurately by manifold descriptions of the boundaries and interfaces.
Both domains are fully fixed along the bottom surface and a distributed load is applied at the top in the $(1,0)$ or $(1,1,0)$ direction for the 2D and the 3D problems, respectively.
The force density $12.5 \times 10^3 \rm{N/mm^2}$ or $12.5 \sqrt{2} \times 10^3 \rm{N/mm^3}$ is applied in 5 steps for the 2D and the 3D problems, respectively.
With respect to the nonlinear solver, the displacement tolerance of the $\mathcal{\ell}_2$ norm for the Newton update is taken to be $10^{-5}$ whereas the relative and absolute tolerances for the residual forces is $10^{-8}$. The relative convergence criteria for the linear solver is $10^{-6}$.
